---
output:
  md_document:
    variant: markdown_github
  html_document:
    variant: markdown_github
    keep_md: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
library(knitr)
opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  fig.align = "center",
  fig.retina = 2,
  out.width = "75%",
  dpi = 96
)
knit_hooks$set(pngquant = hook_pngquant)
```

# backtestPortfolio

Backtest of a portfolio design with real data taken randomly from different
markets and selecting randomly a subset of stocks for proper assessments.
It also allows to backtest multiple portfolio designs and rank them based
on a number of criteria including expected return, Sharpe ratio, drawdown,
etc.


## Installation
```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("dppalomar/backtestPortfolio")

# Getting help
library(backtestPortfolio)
help(package = "backtestPortfolio")
package?backtestPortfolio
?backtestPortfolio
?prices1
```



## Usage of `backtestPortfolio()`
We start by loading the package and some market data from the underlying assets of the S&P 500:
```{r, message = FALSE}
library(backtestPortfolio)
library(xts)
data(prices1)
```
The data `prices1` contains the log-prices of a randomly selected sample of $N=50$ stocks from the
S&P 500 (the package also contains four other sets or randomly selected data, namely, `prices2`, `prices3`, `prices4`, and `prices5`).
```{r}
str(prices1)

colnames(prices1)
```

Now, we define some portfolio design that takes as input the prices and outputs the portfolio vector `w`:
```{r}
# load all the required libraries for the portfolio design
library(CVXR)
library(xts)

naivePortfolioDesign <- function(prices) {
  # compute log returns
  X <- diff(log(prices))[-1]
  
  # compute mean vector and SCM
  mu <- colMeans(X)
  Sigma <- cov(X)
  
  # design mean-variance portfolio
  w <- Variable(nrow(Sigma))
  prob <- Problem(Maximize(t(mu) %*% w - 0.5*quad_form(w, Sigma)),
                  constraints = list(w >= 0, sum(w) == 1))
  result <- solve(prob)
  return(as.vector(result$getValue(w)))
}
```

We are then ready to use the function `backtestPortfolio()` which will execute the portfolio design function on a rolling-based window:
```{r, eval=FALSE}
res <- backtestPortfolio(naivePortfolioDesign, prices1)
print(res)
```



## Links
Package: [GitHub](https://github.com/dppalomar/backtestPortfolio).

README file: [GitHub-readme](https://rawgit.com/dppalomar/backtestPortfolio/master/README.html).
