---
title: "Portfolio Backtesting"
author: |
  | Daniel P. Palomar and Rui ZHOU
  | Hong Kong University of Science and Technology (HKUST)
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    base_format: prettydoc::html_pretty
    theme: tactile
    highlight: vignette
    fig_caption: yes
    number_sections: no
    toc: yes
    toc_depth: 2
  bookdown::pdf_document2:
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 2
header-includes:
  \allowdisplaybreaks
indent: yes
csl: ieee.csl
bibliography: refs.bib
vignette: >
  %\VignetteIndexEntry{Portfolio Backtesting}
  %\VignetteKeyword{portfolio, backtest}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, echo = FALSE}
library(knitr)
opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  fig.retina = 2,
  out.width = "75%",
  dpi = 96
)
knit_hooks$set(pngquant = hook_pngquant)
#Help on bookdown: https://bookdown.org/yihui/bookdown/
#rmarkdown::render("vignettes/PortfolioBacktest-vignette.Rmd", "all")
#rmarkdown::render("vignettes/PortfolioBacktest-vignette.Rmd", "bookdown::html_document2")
#rmarkdown::render("vignettes/PortfolioBacktest-vignette.Rmd", "bookdown::pdf_document2")
#tools::compactPDF("vignettes/PortfolioBacktest-vignette.pdf", gs_quality = "ebook")
```

-----------
> This vignette illustrates the usage of the package `portfolioBacktest` for automated portfolio backtesting.
It can be used by a researcher/practitioner to check a set of different portfolios, as well as by a course 
instructor to evaluate the students in their portfolio design in a fully automated and convenient manner.

# Installation
The package can currently be installed from [GitHub](https://github.com/dppalomar/portfolioBacktest):
```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("dppalomar/portfolioBacktest")

# Getting help
library(portfolioBacktest)
help(package = "portfolioBacktest")
package?portfolioBacktest
?portfolioBacktest
```


# Usage of the package
## Loading data
We start by loading the package and some random sets of stock market data:
```{r, message = FALSE}
library(xts)
library(portfolioBacktest)
data(prices)  # you may want to specify data(prices, package = "portfolioBacktest") 
              # in case there is any conflict with the package PerformanceAnalytics
```
The dataset `prices` is a list of objects `xts` that contains the prices of random sets of stock market data from the S&P 500, HSI, NKY, SHZ, and UKC, over random periods of two years with a random selection of 50 stocks of each universe.
 
```{r}
length(prices)
str(prices[[1]])

colnames(prices[[1]])
```

## Backtesting a single portfolio
We start by defining a simple portfolio design in the form of a function that takes as input the prices and outputs the portfolio vector `w`:
```{r}
uniform_portfolio_fun <- function(prices) {
  N <- ncol(prices)
  w <- rep(1/N, N)  # satisfies the constraints w>=0 amd sum(w)=1
  return(w)
}
```

Now we are ready to use the function `backtestPortfolio()` that will execute and evaluate the portfolio design function on a rolling-window basis:
```{r}
res <- portfolioBacktest(uniform_portfolio_fun, prices[[1]])
names(res)
plot(res$cumPnL)
res$performance
```

Let's try with a slightly more sophisticated portfolio design, like the global minimum variance portfolio (GMVP):
```{r}
GMVP_portfolio_fun <- function(prices) {
  X <- diff(log(prices))[-1]  # compute log returns
  Sigma <- cov(X)  # compute SCM
  # design GMVP
  w <- solve(Sigma, rep(1, nrow(Sigma)))
  w <- w/sum(abs(w))  # it may not satisfy w>=0
  return(w)
}
res <- portfolioBacktest(GMVP_portfolio_fun, prices[[1]])
res$error
res$error_message
```

Indeed, the GMVP does not satisfy the no-shortselling constraint. We can repeat the backtesting indicating that shortselling is allowed:
```{r}
res <- portfolioBacktest(GMVP_portfolio_fun, prices[[1]], shortselling = TRUE)
res$error
res$error_message
res$cpu_time
res$performance
```

We could be more sophisticated and design a Markowitz mean-variance portfolio satisfying the no-shortselling constraint:
```{r, message=FALSE}
library(CVXR)  #install.packages("CVXR")

Markowitz_portfolio_fun <- function(prices) {
  X <- diff(log(prices))[-1]  # compute log returns
  mu <- colMeans(X)  # compute mean vector
  Sigma <- cov(X)  # compute the SCM
  # design mean-variance portfolio
  w <- Variable(nrow(Sigma))
  prob <- Problem(Maximize(t(mu) %*% w - 0.5*quad_form(w, Sigma)),
                  constraints = list(w >= 0, sum(w) == 1))
  result <- solve(prob)
  return(as.vector(result$getValue(w)))
}
```
We can now backtest it:
```{r, cache=TRUE}
res <- portfolioBacktest(Markowitz_portfolio_fun, prices[[1]])
res$error
res$error_message
res$cpu_time
res$performance
```

Instead of backtesting a portfolio on a single `xts` dataset, it is more meaningful to backtest it on multiple datasets. This can be easily done simply by passing a list of `xts` objects:
```{r, cache=TRUE}
res <- portfolioBacktest(Markowitz_portfolio_fun, prices[1:5])
names(res)
res$cpu_time
res$performance
```
In particular, note the additional elements in the returned list:
```{r, cache=TRUE}
res$cpu_time_average
res$performance_summary
res$failure_ratio
```


## Backtesting multiple portfolios
Backtesting multiple portfolios is equally simple. It suffices to pass a list of functions to the backtesting function `multiplePortfolioBacktest()`:
```{r, cache=TRUE}
res <- multiplePortfolioBacktest(portfolio_fun_list = list(uniform_portfolio_fun, 
                                                           GMVP_portfolio_fun), 
                                 prices =  prices[1:5], shortselling = TRUE)
res
```


# Usage for grading students in a course
If an instructor wants to evaluate the students of a course in their portfolio design, it can also be done very easily. It suffices to ask each student to submit a .R script (named `LASTNAME-firstname-STUDENTNUMBER-XXXX.R`) containing the portfolio function called exactly `portfolio_fun()` as well as any other auxiliary functions that it may require (needless to say that the required packages should be loaded in that script with `library()`). Then the instructor can put all those files in a folder and evaluate all of them at once.
```{r, cache=TRUE}
res_all_students <- multiplePortfolioBacktest(folder_path = "folder_path", 
                                              prices =  prices[1:3])
res_all_students$performance_summary
res_all_students$cpu_time_average
res_all_students$failure_ratio
```

Now we can rank the different portfolios/students based on a weighted combination of the rank percentiles (termed scores) of the performance measures:
```{r, fig.height=2, fig.width=9, out.width = "90%"}
leaderboard <- portfolioLeaderboard(res_all_students, weights = list(Sharpe_ratio = 7, max_drawdown = 1, expected_return = 1, ROT = 1))

# show leaderboard
library(gridExtra)
grid.table(leaderboard$leaderboard_scores)
```


## Example of a script file to be submitted by a student
Consider the student Mickey Mouse with id number 666. Then the script file should be named `Mickey-Mouse-666.R` and should contain the portfolio function called exactly `portfolio_fun()` as well as any other auxiliary functions that it may require (needless to say that the required packages should be loaded in that script with `library()`):
```{r, eval=FALSE}
library(CVXR)

auxiliary_function <- function(x) {
  # here whatever code
}

portfolio_fun <- function(prices) {
  X <- diff(log(prices))[-1]  # compute log returns
  mu <- colMeans(X)  # compute mean vector
  Sigma <- cov(X)  # compute the SCM
  # design mean-variance portfolio
  w <- Variable(nrow(Sigma))
  prob <- Problem(Maximize(t(mu) %*% w - 0.5*quad_form(w, Sigma)),
                  constraints = list(w >= 0, sum(w) == 1))
  result <- solve(prob)
  return(as.vector(result$getValue(w)))
}
```






